---
  title: 'Peer-graded Assignment: Statistical Inference Course Project Week 4'
output:
  html_notebook: default
html_document:
  df_print: paged
pdf_document: default
---
  # Statistical Inference Project<br>
  **Author:**Ayush Mohapatra

## Part1:Simulation Exercise <br>

### 1.1 Show the sample mean and compare it to the theoretical mean of the distribution.<br>
This section will execute the simulations and data analysis to illustrate application of the central limit theorem.

```{r}
knitr::opts_chunk$set(echo = TRUE)
lambda <- 0.2
simulation_Data <- matrix(rexp(1000*40, lambda), nrow = 1000, ncol = 40)
distMean <- apply(simulation_Data, 1, mean)
hist(distMean, breaks = 50, main = "The distribution of 1000 averages of 40 random exponentials", xlab = "Value of means", ylab = "Frequency of means", col = "pink")
abline(v = 1/lambda, lty = 1, lwd = 5, col = "red")
legend("topright", lty = 1, lwd = 5, col = "red", legend = "theoretical mean")
```


The simulated sample means are normally distributed .With a center very close to the theoretical mean.<br>
  
  
  ### 1.2 Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.<br>
  
  ```{r}
distVar <- apply(simulation_Data, 1, var)
hist(distVar, breaks = 50, main = "The distribution of 1000 variances of 40 exponentials", xlab = "Value of variances", ylab = "Frequency of variance", col = "light blue")
abline(v = (1/lambda)^2, lty = 1, lwd = 5, col = "blue")
legend("topright", lty = 1, lwd = 5, col = "blue", legend = "theoretical variance")
```

### 1.3 Show that the distribution is approximately normal.<br>

```{r}
par(mfrow = c(3, 1))
hist(simulation_Data, breaks = 50, main = "Distribution of exponentials with lambda equals to 0.2", xlab = "Exponentials", col = "yellow")
hist(distMean, breaks = 50, main = "The distribution of 1000 averages of 40 random exponentials", xlab = "Value of means", ylab = "Frequency of means", col = "pink")
simNorm <- rnorm(1000, mean = mean(distMean), sd = sd(distMean))
hist(simNorm, breaks = 50, main = "A normal distribution with theoretical mean and SD of the exponentials", xlab = "Normal variables", col = "light green")
```

The **first** histogram is the distribution of the exponentials when lambda equals to 0.2. The **second** histogram is the distribution of 1000 averages of 40 random exponentials. The **third** histogram is a real normal distribution with a mean and standard deviation which is equal to the second histogram’s mean and standard deviation .On Comparing the **first** with the **second** histogram, we can see the distrubution becames normal as the means were taken from each groups. It is a result of the central limit theorem. Comparing the **second and the third** histogram, we can see the distribution of the mean is similar to a real normal distribution with the same mean and standard deviation.<br>
  
  # Part 2: Basic Inferential Data Analysis Instructions<br>
  
  ## 2.1 Load the ToothGrowth data and perform some basic exploratory data analyses<br>
  
  ```{r}
library(stats)
data(ToothGrowth)
library(ggplot2)

qplot(dose, len, data = ToothGrowth, color = supp, geom = "point") +  geom_smooth(method = "lm") + labs(title = "ToothGrowth") + labs(x = "Dose of supplements", y = "Length of teeth")
```

### From the plot we can get<br>
1.The length of teeth is proportional to  dose of supplements increases, which indicates that the supplements may help teeth growth.
2.With the same dose, OJ seems to incur a higher increase of teeth growth than VC.
3.the slope of OJ is not as steep as the slope of VC, meaning an increase in VC may make a larger increase in teeth length than in OJ.

## 2.2 Provide a basic summary of the data.<br>
```{r}
summary(ToothGrowth)
```

## 2.3 Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. (Only use the techniques from class, even if there's other approaches worth considering)<br>

If we assume the data is normally distributed, a null hypothesis is indicated and that there is no difference between the mean under each kind of supplements, or each dose of the supplements:
  
  ```{r}
t.test(x = ToothGrowth$len, data = ToothGrowth, paired = FALSE, conf.level = 0.95)$conf.
```

We will be able to construct a confidence interval that 95% of the time, an interval between 16.84 and 20.79 will contain the true mean of the population.

Then we calculate the mean under each kind of supplements, and each dose of the supplements

```{r}
summary(ToothGrowth[ToothGrowth$supp == "OJ", ]$len)[4]
```
```{r}
summary(ToothGrowth[ToothGrowth$supp == "VC", ]$len)[4]
```

Thus the mean of teeth growth after taking OJ is 20.66; the mean of teeth growth after taking VC is 16.96. Both the two are within the confidence interval. We fail to reject the null hypothesis that there is not a difference in teeth growth after taking the two kinds of supplements.

```{r}
summary(ToothGrowth[ToothGrowth$dose == 0.5, ]$len)[4]
```
```{r}
summary(ToothGrowth[ToothGrowth$dose == 1, ]$len)[4]
```
```{r}
summary(ToothGrowth[ToothGrowth$dose == 2, ]$len)[4]
```

Thus the mean of teeth growth after taking dose of 0.5 is 10.6; the mean of teeth growth after taking dose of 1.0 is 19.74; the mean of teeth growth after taking dose of 2.0 is 26.1. We are able to reject the null hypotheis, and there is a difference in teeth growth between each dose of supplements.

Now we know the data is not normally distributed under each dose, along with this conclusion, we may assume the data is normally distributed within each dose. Then we will be able to compare the teeth growth between each supplements under each dose.
```{r}
dose05 <- ToothGrowth[ToothGrowth$dose == 0.5, ]
t.test(x = dose05$len, paired = FALSE, conf.level = 0.95)$conf.
```
```{r}
mean(dose05[dose05$supp == "VC", ]$len)
```
```{r}
mean(dose05[dose05$supp == "OJ", ]$len)
```

Under the dose of 0.5, there are 95% of the time that a confidence interval between 8.50 and 12.71 will contain the true population mean. We also know that the mean of teeth growth after taking 0.5 dose of VC is 7.98 and the mean of teeth growth after taking 0.5 dose of OJ is 13.23. Null hypothesis is rejected.

```{r}
dose10 <- ToothGrowth[ToothGrowth$dose == 1, ]
t.test(x = dose10$len, paired = FALSE, conf.level = 0.95)$conf.
```
```{r}
mean(dose10[dose10$supp == "VC", ]$len)
```
```{r}
mean(dose10[dose10$supp == "OJ", ]$len)
```

Under the dose of 1.0, there are 95% of the time that a confidence interval between 17.67 and 21.80 will contain the true population mean. We also know that the mean of teeth growth after taking 1.0 dose of VC is 16.77 and the mean of teeth growth after taking 1.0 dose of OJ is 22.7. Null hypothesis is rejected.

```{r}
dose20 <- ToothGrowth[ToothGrowth$dose == 2, ]
t.test(x = dose20$len, paired = FALSE, conf.level = 0.95)$conf.
```
```{r}
mean(dose20[dose20$supp == "VC", ]$len)
```

```{r}
mean(dose20[dose20$supp == "OJ", ]$len)
```

Under the dose of 2.0, there are 95% of the time that a confidence interval between 24.33 and 27.87 will contain the true population mean. We also know that the mean of teeth growth after taking 2.0 dose of VC is 26.14 and the mean of teeth growth after taking 2.0 dose of OJ is 26.06. Both of them are within the confidence interval. We fail to reject the null hypothesis.
<br>
  
  ## 2.4 State your conclusions and the assumptions needed for your conclusions.<br>
  When the dose is 0.5 or 1.0 there is a difference between the teeth growth considering OJ and VC. When the dose is 2.0, there is no difference between the teeth growth after taking OJ and VC. The first assumption is that the whole population is normally distributed followed by the assumption that the population is normally distributed under each dose.